---
title: "Taxi_kaggle"
author: "Ariel Erijman"
date: "August 10, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
```

### <font color=blue>__1st step: describe the data__</font> 
```{r echo=FALSE}
train <- read.csv('./train_short.csv')
str(train)
summary(train)
```

### <font color=blue>__Structure and Summary of random 100000 samples from the dataset__</font>
1. 1 id per sample
2. 2 vendors with fairly equally distributed number of trips
3. pickup and dropoff datetimes
4. Number of passengers per trip
5. Pickup and dripoff x,y locations
6. Store and fwd flag ???
7. Trip duration has a median = 11min and mean = 15min. The min(1min) and max(23hs) don't make much sense to me...

```{r echo = FALSE}
# Date & time to integers
train$hour <- as.POSIXlt(train$pickup_datetime)$hour
train$day <- as.POSIXlt(train$pickup_datetime)$mday
train$month <- as.POSIXlt(train$pickup_datetime)$mon
train$weekday <- as.POSIXlt(train$pickup_datetime)$wday
```

### <font color=blue>__Distribution of trip duration__</font>
```{r echo=FALSE}
hist(train$trip_duration[train$trip_duration<=6000], breaks = 100, main = "Trip duration")
```

### <font color=blue>__What are the most common hours for a trip in NY?__</font>
It looks clear that mostf the trips are between ~9AM and ~19PM.
```{r echo=FALSE}
boxplot(train$hour, main="Hours stats for a trip in NY?")
```

### <font color=blue>__How does the duration varies with the hour of the day?__</font>
The single significant difference in trip duration influenced by time of the day is @5&6 AM
```{r echo=FALSE}
# define a subset of train with lower values of trip duration to overcome the vast ammount of outlyers
train_less <- train[train$trip_duration < 1000,]
boxplot(
  train_less$trip_duration~train_less$hour, 
  xlab='hour',
  ylab='trip duration',
  main="Trip duration by hours"
  )
```

### <font color=blue>__Does the # of passengers has something to do with trip duration?__</font>
Should we include 0 passengers trips?
>5 passengers probably includes a bigger car. Nevertheless, it does not affect the trip duration at all
```{r echo=FALSE}
boxplot(train_less$trip_duration~train_less$passenger_count, main="Trip duration by passenger count")
```
